In-silico molecular enrichment and clearance of the human intracranial space

### 2024-12-16 paper/supplementary_main.tex

The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
space averaged over each perivascular cross-section and is governed by 
\begin{equation}
  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s (\hat c) ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
\label{eq:1d_pde}
\end{equation}
where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
associated PVS $\Omega^i$, the area of the cross-section
$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
and $\hat{u}$ is a convective velocity representing an (average) CSF
flow velocity in the axial direction of the perivascular spaces.

@@

To discretize~\eqref{eq:1d_pde} with the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
linear polynomials defined over $\Lambda$:
\begin{equation}
  \hat V_h
  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
\end{equation}
where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$. The discrete formulation then reads: for $n = 1,2,\ldots, $ find $\hat c_h^n \in \hat V_h$ such that for all $\hat v \in \hat V_h^k$: 
\begin{align}\label{eq:1d_dg_disc}
\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.  
\end{align}
 In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion stabilization term, for more details see for e.g. \cite[Section 12.6]{quarteroni2009numerical}: 
 \begin{align}
    a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) =  \frac{1}{\|\hat u\|_{L^2(\Lambda^i)}} 
 h_{\Lambda^i} \int_{\Lambda^i} A (\hat u \partial_s \hat c_h  ) (\hat u \partial_s \hat v),  
 \end{align}
 where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.

 \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} comlemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0$ is the following:

>>>>

The concentration $\hat{c} = \hat{c}(s, t)$ in the periarterial and
perivenous networks ($s \in \Lambda_a, s \in \Lambda_v$), entering
in~\eqref{eq:3d_pde}, represents the concentration in the perivascular
space averaged over each perivascular cross-section and is governed by 
\begin{equation}
  \partial_t (A  \hat c) - \partial_s( \hat D A \partial_s \hat c  ) + \partial_s(A \hat u \hat c )  +  \xi P (\hat c - \overline{c}) = 0
  \qquad \text{in} \quad \Lambda_a, \Lambda_v ,
\label{eq:1d_pde}
\end{equation}
where in $\Lambda_a, \Lambda_v$ refers to in each $\Lambda^i$ in each
of these networks. In~\eqref{eq:1d_pde}, $A = A(s)$ is defined as the
PVS cross-sectional area; i.e.,~for each $s \in \Lambda^i$ with
associated PVS $\Omega^i$, the area of the cross-section
$\Theta^i(s)$. Also, $\hat{D}$ is the effective diffusion coefficient,
and $\hat{u}$ is a convective velocity representing an (average) CSF
flow velocity in the axial direction of the perivascular spaces.

@@

To discretize~\eqref{eq:1d_pde} with the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:1d_bcs}, we use the space of continuous piecewise
linear polynomials defined over $\Lambda$:
\begin{equation}
  \hat V_h
  = \{ v \in C^0(\Lambda), \; v \vert_{\Lambda^i} \in \mathrm{P}^1(\Lambda^i) \}, 
\end{equation}
where $\mathrm{P}^1(\Lambda^i)$ is the space of linear polynomials on each $\Lambda^i$. The discrete formulation then reads: for $n = 1,2,\ldots, $ find $\hat c_h^n \in \hat V_h$ such that for all $\hat v \in \hat V_h^k$: 
\begin{align}\label{eq:1d_dg_disc}
\int_{\Lambda}  \frac{1}{\tau}(A  \hat c_h^n - A \hat c_h^{n-1}) \hat v  + \int_{\Lambda} \hat D A \partial_s \hat c_h^n  \partial_s \hat v^n - \int_{\Lambda} A \hat u \hat c_h^n \partial_s \hat v   +  \int_{\Lambda} \xi P (\hat c_h^n - \overline{c_h^n}) \hat v +a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) = 0.  
\end{align}
 In the above, if $\hat u$ is non--zero on $\Lambda^i$, then the artificial diffusion $a_h^{\mathrm{stab}}$ stabilization term is nonzero  and is given below, for more details see for  \cite[Section 12.6]{quarteroni2009numerical}. 
 \begin{align}
    a_h^{\mathrm{stab}}(\hat u; \hat c, \hat v) =  \frac{1}{\|\hat u\|_{L^2(\Lambda^i)}} 
 h_{\Lambda^i} \int_{\Lambda^i} A (\hat u \partial_s \hat c_h  ) (\hat u \partial_s \hat v),  
 \end{align}
 where $h_{\Lambda^i}$ is the mesh--size of $\Lambda^i$.

 \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} comlemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:
